---
title: "lab6"
author: "Thomas Lenihan"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Equal ranks (or medians)
HA: Ranks are not equal (or medians are not equal)

###Create samples

`set.seed()` to create "pseudorandom, but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```


```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results: 

```{r}
ttest <- t.test(gp_1, gp_2)
#If these data were paired, I can just add he argument `paired = "TRUE"` within the t.test() function. 
```

This p-value means that IF our samples are drawn from populations with the same mean (i.e., if the null hypothesis is true), there is `r round(ttest$p.value,2)` probability of finding sample means *at least as different* as ours by random chance. 

[A more comprehensive explanation of difference between groups...actually means, data spread, effect size]...there is no significant difference (t(df) = t-statistic value, p = ).

### Mann Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that IF our samples are drawn from populations with the same median (i.e., if the null hypothesis is true), there is `r round(mwu$p.value,2)` probability of finding sample means *at least as different* as ours by random chance. 

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value,2)`).

Example text for reporting: "The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly (Mann Whitney U,...)"

## Simple linear regression

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense. 

Use the `penguins` data from (palmerpenguins), exploring flipper length (x) versus body mass (y).

```{r}
# Use View(penguins) to look at it
# Make an exploratory scatter plot of flipper length (x-axis) versus body mass (y-axis)
ggplot (data = penguins, 
        aes(x = flipper_length_mm,
            y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
  
```

### Find the model estimates (coefficients)

Use the `lm()` function for  linear model. 

body_mass_g ~ <<<< tilde (not a minus sign)
- tilde = "as a function of"

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

Model Coefficients: `r penguin_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients[2]`

### Use `broom::tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy

penguin_lm_tidy$estimate[2]
```

### Use `broom::glance()` to get output of model fit & statistical significance

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use `equatiomatic::extract_eq()` to return the model in LaTeX

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use `plot()` to automatically produce diagnostic plots

```{r}
plot(penguin_lm)
```

